### 2022-04-02 outputs/paper.Rmd
some grammar corrections

---
title: "Childcare while their mom is working"
author: 
  - Pascal Lee Slew
  - Yunkyung Park
thanks: "Code and data are available at: https://github.com/Pascal-304/dhs_analysis."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "Working mothers are often faced with the conflict of responsibility towards their children and the role they play at work. The Ghana Demographic and Health Survey (GDHS) forms part of a global initiative to collect data on family, fertility, and child health. In this paper, we used the 1998 GDHS and focused on the child care and working mothers aspect. Demographics and other aspects such as education tend to predict the child care preference. We found that families living in Greater Accra are more likely to send their child to school and the majority of working mothers look after their child by themselves. The analysis of the child care and working mothers data provides an insight on current trends and has significant implications for public policy and program decisions."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library, echo=FALSE, message=FALSE, warning = FALSE}
# install.packages('bookdown')
# install.packages("stringr")
# install.packages("ggforce")
library(stringr) 
library(dplyr)
library(tidyverse) 
library(janitor)
library(ggplot2)
library(knitr)
library(kableExtra)
library(bookdown)
library(ggforce)
```

```{r readdata, echo=FALSE}
ghana <- read_csv('data/cleaned_data.csv', show_col_types = FALSE)
```

# Introduction

Child care matter to the development of a child. Previously, women were thought to fit the traditional role of a mother, being at home and looking after her children. This thought is being challenged with a growing presence of women in the workforce, contributing to economic activity. Women are often faced with the conflicting role they have at home and at work. However, sometimes, mothers have to work out of necessity to contribute to the financial situation of their families. Families have to find ways to raise their children. The quality of child care given tend to have positive effects on the development of the child [@care]. Having an idea of the current child care practices is important since it gives crucial information on the weaknesses of current policies and would give an insight on how to improve the situation around child care.

In this paper, we examined child care decisions of working mothers across regions, the mothers' highest education level attained, their occupation, work and employment status in Ghana. We conducted a descriptive analysis of child status and childcare preferences. We first looked at the number of working mothers from the sample population, then examined child care practices based on employment status, education level among other factors. The differences in child care practices were noticeable in different regions. We found that most working mothers tend to look after their young child by themselves and those living in Greater Accra have a higher tendency to put their child in child care centres compared to other regions. Another observation was the highest level of education attained may have an influence on the number of children they have. The proportion of working mothers with at least secondary education with no child under six years old is lower than those with at least one or more child under six.

The rest of the paper is divided into 4 sections. Section \@ref(data) explains the data source, the survey methodology and some of its key features, strengths and weaknesses. Section \@ref(results) relays the findings through graphs. Section @\ref(discussion) discusses in more detail what was found and some limitations of what we did and Section \@ref(appendix) contains a datasheet for the dataset we used.

# Data

## Data Source

This paper uses the Childcare While Working Data [@ghana1998] obtained from the 1998 Ghana Demographic and Health Survey (GDHS). The dataset was collected by downloading the final report of the 1998 GDHS published in the Demographic and Health Surveys (DHS) programme. The package 'pdftools' [@citepdftools] is used to download and read the content into R by parsing the pdf. The package 'stringi' [@citestringi] and 'tidyverse' [@citetidyverse] are then used to gather the data.

## Methodology

The 1998 Ghana Demographic and Health Survey was a national-level survey carried out in Ghana. This was a part of the worldwide Demographic and Health Surveys (DHS) programme and was conducted by the Ghana Statistical Service in collaboration with the Macro International Inc. It was designed to collect data on fertility, family planning, and maternal and child health. The 1998 GDHS was conducted from mid-November 1998 to mid-February 1999.

The survey frame was created using the list of Enumeration Areas (EAs) with population and household information from the 1984 Population Census. The 1998 GDHS used a two-stage stratified sampling method. At the first stage of sampling, 400 EAs were selected using systematic sampling with probability proportional to size where the selected EAs comprised 138 in the urban areas and 262 in the rural areas. A complete household listing operation was then carried out in all the selected EAs to provide a sampling frame for the second stage selection of households. At the second stage of sampling, a systematic sample of 20 households per EA was selected in the Northern, Upper West and Upper East Regions, and 15 households per EA was selected in the other regions.

The target sample size for the 1998 GDHS was 6,000 consisted of 4,500 women of age 15-49 and men of age 15-59. A total of 4,843 women age 15-49 and 1,546 men age 15-59 completed the survey. To account for non-response, 6375 households were selected for interview.

The GDHS was composed of 3 parts: the Household's questionnaire, the Woman's questionnaire and the Man's questionnaire. All the questionnaires were modified to account for the situation in Ghana.

The survey was conducted in person. The interview staff were first trained for a period of three weeks. It was conducted by a total of 14 teams, comprising a team supervisor, a field editor, three interviewers and a driver. The response rate was 97% for individual men and women selected. Non-response was mainly due to inability to find the selected interviewee at home. Repeated callbacks were made to try to reach the selected respondents. The results of the survey did not disclose any confidential information that might identify the respondents. The survey was a voluntary one but it is unclear if the purpose of the survey explained to the respondents for them to make an informed decision.

## Strengths

The questionnaires were made available into the five major local languages (Akan, Ga, Ewe, Hausa, and Dagbani), making the survey accessible to a larger population. After the household listing, sport checks were conducted to make sure the work has been done correctly. Before the actual survey, a test was conducted and adjustments were made to the questionnaires based on the feedback obtained from the test. The questionnaire questions were comprehensive and detailed

## Weaknesses

Since the questionnaires were only available in the five major local languages, if the selected household was from a minor ethnic group and know only their own language, then it would be difficult for the interviewee to understand and answer the survey. Visitors who stayed a night before in the selected household were eligible for the interview. The visitor may be from a different region, so the targeted sample size for each region may be skewed.

## Key features

```{r data, echo=FALSE}
ghana %>%
  select(background, no_child_under_six, one_more_child_under_six, caretaker_respondent, number_of_employed_women, character) %>%
  relocate(no_child_under_six, .after = one_more_child_under_six) %>%
  slice(5:6, 13:20, 26) %>%
  kable(caption='A subset of key features',
        col.names = c('Background Characteristic', 'Yes', 'No', 'Respondent', 'Number of employed women', 'Category'),
        digits = 1,
        booktabs = TRUE,
        linesep = "",
        align = c('l', 'c', 'c', 'c', 'c', 'l'),) %>%
  add_header_above(c(" " = 1, "Child under six" = 2, "Child's caretaker" = 1, " " = 2),
                   escape=F) %>%
  add_footnote("Decimals are in percentage (%).")
```

Some key features include background and a percent distribution of currently employed women by whether they have a child under six years of age at home according to selected background characteristics, Ghana 1998. If the respondent answered to have one or more child under six, then they were further asked to indicate the caretaker of the child such as the respondent themselves, their partner, or whether they hired a person for it, etc.

The dataset is processed and analyzed using 'R' [@citeR] mainly with the 'tidyverse' [@citetidyverse] and 'dplyr' [@citedplyr] packages. The 'janitor' [@citejanitor] is used to clean data, and the graphs are created in 'ggplot2' [@citeggplot2] and 'ggforce' [@citeggforce] packages. The map is created using in 'rnaturalearth' [@citernaturalearth] and 'sf' [@citesf] packages. The package 'stringr' [@citestringr] is used to manipulate the character string. The packages 'bookdown' [@citebookdown] and 'knitr' [@citeknitr] are used in generating the R Markdown report. Table \@ref(tab:data) created using 'kableExtra' [@citekableExtra] shows a subset of key features that will be discussed in this paper.

## General Overview

```{r generalchildcare, echo=FALSE, fig.cap="Distribution of the caretaker of the child when the mother is working", fig.width=6, fig.height=4}
totl <-
  ghana %>%
  filter(character=="Total") %>%
  mutate(
    no_child_under_six = round(no_child_under_six/100 * number_of_employed_women, 0),
    one_more_child_under_six = round(one_more_child_under_six/100 * number_of_employed_women, 0)
  )
# make dataset in long format
totl_long <-
  totl %>%
  select(-c(character, total)) %>%
  pivot_longer(!background, names_to = "variable", values_to = "count")

# caretaker aspect
totl_caretaker <-
  totl_long %>%
  filter(str_detect(variable, "^caretaker_")) %>%
  mutate(variable = str_replace_all(variable, "caretaker_", "")) %>%
  rename(caretaker = variable,
         percentage = count) %>%
  mutate(
    caretaker = factor(caretaker, 
                       levels = c('respondent',
                                  'partner',
                                  'other_relative',
                                  'neighbor',
                                  'hired',
                                  'child_in_school',
                                  'other_female_child',
                                  'other_male_child',
                                  'not_worked_since_birth',
                                  'other',
                                  'missing')
    )
  )

# plot the bar graph
totl_caretaker %>%
  ggplot(mapping=aes(x=caretaker, y=percentage, fill=caretaker)) +
  geom_bar(stat = "identity", position="dodge") +
  geom_text(aes(label=paste(percentage, '%'), y=percentage+2),
            size = 3,
            position = position_dodge(width=0.5)) +
  theme_minimal() +
  labs(x='') +
  theme(legend.position='none') +
  coord_flip() + 
  scale_x_discrete(
    labels = 
      c("respondent" = "respondent",
        "partner" = "partner",
        "other_relative" = "relative",
        "neighbor" = "neighbor",
        "hired" = "hired",
        "child_in_school" = "schooling",
        "other_female_child" = "other female child",
        "other_male_child" = "other male child",
        "not_worked_since_birth" = "not worked since birth",
        "other" = "other",
        "missing" = "missing")
  )
```

We can tell from Table \@ref(tab:data) that there are 3,564 working mothers, and over a half of them have a child under six years. Figure \@ref(fig:generalchildcare) illustrates that 47.6% of these mothers look after their own children while they are at work, 22.1% have their relatives other than their husband/partner to take care of the child, 14% send the child to school or other institutional care. On top of that, less than 3 percent of women have husbands/partners to look after the child while they are at work.

It should be noted that 1.9% has not worked since the last birth, and this indicates the percentage of the respondents who were not working at the time of the survey but had been employed at some time during the last 12 months.

# Results

## Region

```{r map, message=FALSE, echo = FALSE, warning=FALSE, fig.cap="Ghana administrative regions"}
# install.packages('rnaturalearth')
# install.packages('sf')
library(rnaturalearth) # we need to cite these packages as well
library(sf)
# devtools::install_github("ropensci/rnaturalearthhires")
# The st_centroid function takes the sf object and returns the center of 
# the polygon, again as a sf object.
map_ghana <- ne_states(country = 'Ghana', 
                       returnclass = 'sf')  # grab provinces within Ghana

map_ghana %>% 
  ggplot() +
  geom_sf( ) + 
  geom_text(data = st_centroid(map_ghana), 
            size = 3,
             aes(x = longitude, 
                 y = latitude,
                 label= woe_name)) +
  theme_minimal()
```

Figure \@ref(fig:map) shows a map of Ghana and its administrative regions. Ghana has 10 administrative regions, namely, Upper West, Upper East, Northern, Brong Ahato, Western, Ashanti, Central, Eastern, Volta and Greater Accra. In 1998, about 57.6% of the Ghanese population were living in rural areas [@population]. The Greater Accra is considered the most urbanized region of Ghana.

```{r residence, echo=FALSE, fig.cap="Comparison of the number of working mothers between urban and rural areas"}
# Urban vs Rural

# Obtain subset of dataset for residence area
residence <- ghana %>% 
  filter(character == 'Residence')

# look at survey respondents distribution in urban and rural areas
residence %>% 
  ggplot(aes(x = background, y = number_of_employed_women)) +
  geom_bar(stat = 'identity', fill = "burlywood") +
  labs(x = "Area",
       y = "Number of employed women") +
  theme_minimal()
```

Figure \@ref(fig:residence) shows the number of employed mothers in urban and rural areas. There are 1127 women who were selected in Urban areas while 2338 mothers were selected in rural areas. There are less women living in urban who are currently working, and this is not surprising as a larger population live in rural areas.

```{r region, echo=FALSE, fig.cap = "Comparison of the number of working mothers selected in the different administrative regions"}
# region subdataset
region <- ghana %>% 
  filter(character == "Region") %>% 
  mutate(background = recode(background,
                             'Western' = 'Western',
                             'Central' = 'Central',
                             'Greater_Accra' = 'Greater Accra',
                             'Volta' = 'Volta',
                             'Eastern' = 'Eastern',
                             'Ashanti' = 'Ashanti',
                             'Brong_Ahafo' = 'Brong Ahafo',
                             'Northen' = 'Northern',
                             'Upper_West' = 'Upper West',
                             'Upper_East' = 'Upper East'))

# number of survey respondents in the different administrative regions
region %>% 
  ggplot(aes(x = background,
             y = number_of_employed_women)) +
  geom_bar(stat = 'identity',
           fill = 'burlywood') +
  labs(x = "Administrative regions",
       y = "Number of employed women") +
  theme_minimal() +
  coord_flip()
```

Figure @\ref(fig:region) shows the number of employed mothers in the different administrative regions. There were more working mothers in Greater Accra and Ashanti while Upper West had the least number of working mothers. The reason might be because of the number of people living in the different administrative regions are different, therefore, the sample size differed from one region to another. Greater Accra is the most urbanized region which implies more work opportunities and therefore, a larger population would be concentrated there.

```{r childstatus, echo = FALSE, fig.cap = "Comparing the percentage of mothers having no child under six to one or more in both Urban and Rural areas"}
# make dataset in long format
resid_long <- residence %>% 
  select(-c(character)) %>% 
  pivot_longer(!background, names_to = "variables", values_to = "percentage") %>% 
  mutate(variables = recode(variables,
                            'no_child_under_six' = "No child under six", 
                            'one_more_child_under_six' = "One or more child under six",
                            'caretaker_respondent' = 'respondent',
                             'caretaker_partner' = 'partner',
                             'caretaker_other_relative' = 'other relative',
                             'caretaker_other_male_child' = 'other male child',
                             'caretaker_other_female_child' = 'other female child',
                             'caretaker_other' = 'other',
                             'caretaker_not_worked since_birth' = 'not worked since birth',
                             'caretaker_neighbor' = 'neighbour',
                             'caretaker_missing' = 'missing',
                             'caretaker_hired' = 'hired',
                             'caretaker_child_in_school' = 'child in school'))

# create label names list
label_list <- list('no_child_under_six' = "No child under six", 
                   'one_more_child_under_six' = "One or more child under six")

labellers <- function(variable, value) {
  return(label_list[value])
}

# plot bar graph
resid_long %>% 
  filter(variables == 'No child under six' | variables == 'One or more child under six') %>% 
  ggplot(aes(x = background, 
             y = percentage,
             fill = variables)) +
  geom_bar(stat = 'identity',
           position = 'dodge') +
  labs(x = "Area",
       y = "Percent",
       fill = "Child status") +
  theme_minimal() +
  scale_fill_brewer() +
  theme(legend.position = "bottom")

```

Figure \@ref(fig:childstatus) shows the child status of families in Urban and Rural areas. In rural areas, there is a larger percentage of working mothers having one or more child under six years old. The converse is true in Urban areas. Working mothers tend to have no child under six years old.

```{r areachildcare, echo=FALSE, fig.cap = "Comparing child caretaker preferences in Urban and Rural areas"}
# caretaker aspect
# make dataset into long format
# rename 
caretaker <- residence %>% 
  select(-c(character, no_child_under_six, one_more_child_under_six, total, number_of_employed_women, caretaker_missing)) %>% 
  pivot_longer(!background, names_to = "caretakers", values_to = "percentage") %>% 
  mutate(caretakers = recode(caretakers,
                             'caretaker_respondent' = 'respondent',
                             'caretaker_partner' = 'partner',
                             'caretaker_other_relative' = 'other relative',
                             'caretaker_other_male_child' = 'other male child',
                             'caretaker_other_female_child' = 'other female child',
                             'caretaker_other' = 'other',
                             'caretaker_not_worked_since_birth' = 'not worked since birth',
                             'caretaker_neighbor' = 'neighbour',
                             'caretaker_missing' = 'missing',
                             'caretaker_hired' = 'hired',
                             'caretaker_child_in_school' = 'child in school'))

# plot bar graph
caretaker %>% 
  ggplot(aes(x = caretakers, 
             y = percentage,
             fill = background)) +
  geom_bar(stat = 'identity',
           position = "dodge") +
  labs(x = "Caretaker",
       y = "Percentage of working mothers",
       fill = "Area") +
  theme_minimal() +
  coord_flip() +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom")

```

Figure @\ref(fig:areachildcare) shows the Ghanese families'caretaking preferences in both urban and rural areas. The preferred caretaking method in both urban and rural areas is the working mothers looking after their children by themselves. The next preferred method for in both areas would be to let the children be looked after by relatives. Other notable observations would be that families living in urban areas tend to put their child in school and in rural areas, families let the older female child to care for the under six years old child much more than in urban areas.

```{r caretakingregions, echo = FALSE, fig.cap = "Caretaking behaviour in the different administrative regions", fig.width=7, fig.height=7}
# make dataset into long format
# recode names
child <- region %>% 
  select(-c(character, total, caretaker_missing)) %>% 
  pivot_longer(!background, names_to = "x", values_to = "percentage") %>% 
  mutate(x = recode(x,
                    'caretaker_respondent' = 'respondent',
                    'caretaker_partner' = 'partner',
                    'caretaker_other_relative' = 'other relative',
                    'caretaker_other_male_child' = 'other male child',
                    'caretaker_other_female_child' = 'other female child',
                    'caretaker_other' = 'other',
                    'caretaker_not_worked_since_birth' = 'not worked since birth',
                    'caretaker_neighbor' = 'neighbour',
                    'caretaker_missing' = 'missing',
                    'caretaker_hired' = 'hired',
                    'caretaker_child_in_school' = 'child in school',
                    'no_child_under_six' = 'no child under six',
                    'one_more_child_under_six' = 'one or more child under six',
                    'number_of_employed_women' = 'number of employed women'))

# plot bar graph for caretaker comparisons and administrative regions
child %>% 
  filter(x != 'number of employed women', x != 'no child under six', x!= 'one or more child under six') %>% 
  ggplot(aes(x = x,
             y = percentage,
             fill = background)) +
  geom_bar(stat = 'identity') + 
  labs(x = "caretaking preference",
       y = "percent") +
  theme_minimal() +
  coord_flip() +
  facet_wrap(~background) +
  theme(legend.position = "none")

```

Figure \@ref(fig:caretakingregions) shows the caretaking preference of Ghanese families in the different administrative regions. Working mothers tend to look after their child themselves the most in all regions. In Greater Accra, families choose to put their child in school the most compared to other administrative regions. Another observation is that the proportions of families opting to put their child in school in Northern, Upper East and Upper West are the lowest, namely 1.3%, 2.7% and 2.0% respectively. The other popular caretaking options after by the mother herself are by a relative or by another older female child.

```{r chstatus, echo = FALSE, fig.cap = "Comparing the child status in different administrative regions"}
# looking at the number of employed women selected in the different administrative regions
# with the child under six
child %>% 
  filter(x == "no child under six" | x == "one or more child under six") %>% 
  ggplot(aes(x = background,
             y = percentage,
             fill = x)) +
  geom_bar(stat = 'identity', position = "dodge", width = 0.8) +
  labs(x = "child status",
       y = "percentage",
       fill = "child status") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  coord_flip()

```

Figure \@ref(fig:chstatus) shows the child status of Ghanese families in the different administrative regions. By child status, we mean the whether the families have children under six years old. Ghanese families tend to have one or more child under six years old more than no child under six with the exception of Greater Accra. The proportion of families with one or more child under six is slightly more pronounced in Northern, Upper West, Brong Ahafo and Upper East regions.

## Mother's education

```{r motheredu, echo=FALSE, fig.cap="Mother's Education level", fig.width=6, fig.height=4}
# select mother edu
mother_edu <-
  ghana %>%
  filter(character=="Mother\'s_education") %>%
  mutate(
    no_child_under_six = round(no_child_under_six/100 * number_of_employed_women, 0),
    one_more_child_under_six = round(one_more_child_under_six/100 * number_of_employed_women, 0)
  )

# make dataset in long format
edu_long <-
  mother_edu %>%
  select(-c(character, total)) %>%
  pivot_longer(!background, names_to = "variable", values_to = "count") %>%
  mutate(
    background = factor(background,
                        levels = c('No_education',
                                   'Primary',
                                   'Middle',
                                   'Secondary+')
    )
  )

edu_long %>%
  filter(variable == 'no_child_under_six' | variable == 'one_more_child_under_six') %>%
  ggplot(aes(x = background, y = count, fill = variable)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Mother's education", y = "Number of employed women", fill = "Child under six") +
  theme_minimal() +
  scale_fill_discrete(labels = c("No child under six", 
                                 "One or more child under six")) +
  scale_x_discrete(labels = c("No_education" = "No education")) +
  theme(legend.position = 'bottom')
```

Figure \@ref(fig:motheredu) shows that there are significantly less employed women who studied at secondary school or higher. Also, it can be observed that the working mothers who did not continue their education until the secondary school are more likely to have a child under six, whereas there are more employed women who studied beyond the secondary school and have no child under six. In addition, when we take a look at the number of employed women who did not get any education, we can see that there are a lot more women with a child under six than without, and the difference between them is notably big. However, as the level of working mother's education gets higher, the gap between having a child under six and no child under six decreases, and it gets reversed when the women completed the secondary school.

```{r mothereduchildcare, echo=FALSE, fig.cap="Distribution of child's caretaker by mother's education level", fig.width=6, fig.height=4}
# caretaker aspect
edu_caretaker <-
  edu_long %>%
  filter(str_detect(variable, "^caretaker_")) %>%
  mutate(variable = str_replace_all(variable, "caretaker_", "")) %>%
  rename(caretaker = variable,
         percentage = count) %>%
  mutate(
    caretaker = factor(caretaker, 
                       levels = c('respondent',
                                  'partner',
                                  'other_relative',
                                  'neighbor',
                                  'hired',
                                  'child_in_school',
                                  'other_female_child',
                                  'other_male_child',
                                  'not_worked_since_birth',
                                  'other',
                                  'missing')
    )
  )

# plot the bar graph
edu_caretaker %>%
  ggplot(mapping=aes(x=caretaker, y=percentage, fill=caretaker)) +
  geom_bar(stat = "identity", position="dodge") + 
  theme_minimal() +
  labs(x='Caretaker while at work', y="Percentage (in %)") +
  theme(legend.position='none') +
  coord_flip() + 
  scale_x_discrete(
    labels = 
      c("respondent" = "respondent",
        "partner" = "partner",
        "other_relative" = "relative",
        "neighbor" = "neighbor",
        "hired" = "hired",
        "child_in_school" = "schooling",
        "other_female_child" = "other female child",
        "other_male_child" = "other male child",
        "not_worked_since_birth" = "not worked since birth",
        "other" = "other",
        "missing" = "missing")
  ) +
  facet_wrap(vars(background),
             dir = "v",
             scales = "free",
             labeller = labeller(background = 
                                   c("No_education" = "No education",
                                     "Primary" = "Primary",
                                     "Middle" = "Middle",
                                     "Secondary+" = "Secondary+")
                                 )
  )
```

Figure \@ref(fig:mothereduchildcare) demonstrates that no matter what their education level is, it is mostly the employed women who take care of their children while working. When the level of education is lower, then it is more likely for the respondent to look after their own children under six years. Over 50% of the women with no education take care of their own children whereas less than 40% of the women with secondary or higher levels of education do. As the level of the education gets higher, then the percentage of families choosing the option of schooling their children gets higher as well. Moreover, it is worthwhile to mention that the women who completed high school are more likely to stop working after having a birth compared to the women whose education level is lower than secondary school.

## Work status

```{r workstatus, echo=FALSE, message=FALSE, fig.cap="Work status", fig.width=6, fig.height=4}
# select work status
work_status <-
  ghana %>%
  filter(character=="Work_status") %>%
  mutate(
    no_child_under_six = round(no_child_under_six/100 * number_of_employed_women, 0),
    one_more_child_under_six = round(one_more_child_under_six/100 * number_of_employed_women, 0)
  )

# make dataset in long format
work_long <-
  work_status %>%
  select(-c(character, total)) %>%
  pivot_longer(!background, names_to = "variable", values_to = "count") %>%
  mutate(
    background = factor(background,
                        levels = c('For_family_member',
                                   'For_someone_else',
                                   'Self-employed')
    )
  )

work_long %>%
  filter(variable == 'no_child_under_six' | variable == 'one_more_child_under_six') %>%
  ggplot(aes(x = background, y = count, fill = variable)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Work status", y = "Number of employed women", fill = "Child under six") +
  theme_minimal() +
  scale_x_discrete(labels = c("For_family_member" = "For family member",
                              "For_someone_else" = "For someone else")) + 
  scale_fill_discrete(labels = c("No child under six", 
                                 "One or more child under six")) +
  theme(legend.position='bottom')
```

Figure \@ref(fig:workstatus) shows that most employed mothers are self-employed. If the women work for themselves or for their family, then it is more likely to have one or more than one child under six years old, whereas when they work for someone else, there are more women who do not have a child under six years old. It should be noted that no child under six does not mean that they have no child.

```{r workstatuschildcare, echo=FALSE, fig.cap="Distribution of child's caretaker by work status", fig.width=6, fig.height=4}
# caretaker aspect
work_caretaker <-
  work_long %>%
  filter(str_detect(variable, "^caretaker_")) %>%
  mutate(variable = str_replace_all(variable, "caretaker_", "")) %>%
  rename(caretaker = variable,
         percentage = count) %>%
  mutate(
    caretaker = factor(caretaker, 
                       levels = c('respondent',
                                  'partner',
                                  'other_relative',
                                  'neighbor',
                                  'hired',
                                  'child_in_school',
                                  'other_female_child',
                                  'other_male_child',
                                  'not_worked_since_birth',
                                  'other',
                                  'missing')
    )
  )

# plot the bar graph
work_caretaker %>%
  ggplot(mapping=aes(x=caretaker, y=percentage, fill=caretaker)) +
  geom_bar(stat = "identity", position="dodge") + 
  theme_minimal() +
  labs(x='Caretaker while at work', y="Percent") +
  theme(legend.position='none') +
  coord_flip() + 
  scale_x_discrete(
    labels = 
      c("respondent" = "respondent",
        "partner" = "partner",
        "other_relative" = "relative",
        "neighbor" = "neighbor",
        "hired" = "hired",
        "child_in_school" = "schooling",
        "other_female_child" = "other female child",
        "other_male_child" = "other male child",
        "not_worked_since_birth" = "not worked since birth",
        "other" = "other",
        "missing" = "missing")
  ) +
  facet_wrap_paginate(vars(background),
                      ncol=2,
                      nrow=2,
                      dir = "h",
                      scales = "free",
                      labeller = labeller(background = c("For_family_member" = "For family member",
                                                         "For_someone_else" = "For someone else",
                                                         "Self-employed" = "Self-employed")
                      )
  )

```

It can be observed from Figure \@ref(fig:workstatuschildcare) that most employed women take care of their children by themselves regardless of whom they work for. It is their relatives who take care of their children while they are at work. We can tell that if they are self-employed or work for the family, then it is more likely to have a family member taking care of their children under six years old, whereas the percentage of families opting for schooling option is relatively higher when they work for someone else.

## Occupation

```{r occupation, echo=FALSE, message=FALSE, fig.cap="Occupation", fig.width=6, fig.height=4}
# select occupation
occupation <-
  ghana %>%
  filter(character=="Occupation") %>%
  mutate(
    no_child_under_six = round(no_child_under_six/100 * number_of_employed_women, 0),
    one_more_child_under_six = round(one_more_child_under_six/100 * number_of_employed_women, 0)
  )

# make dataset in long format
occup_long <-
  occupation %>%
  select(-c(character, total)) %>%
  pivot_longer(!background, names_to = "variable", values_to = "count") %>%
  mutate(
    background = factor(background,
                        levels = c('Agricultural',
                                   'Nonagricultural')
    )
  )

occup_long %>%
  filter(variable == 'no_child_under_six' | variable == 'one_more_child_under_six') %>%
  ggplot(aes(x = background, 
             y = count,
             fill = variable)) +
  geom_bar(stat = "identity",
           position = "dodge") +
  labs(x = "Sector",
       y = "Number of employed women",
       fill = "Child under six") +
  theme_minimal() +
  scale_fill_discrete(labels = c("No child under six", 
                                 "One or more child under six")) +
  theme(legend.position = 'bottom')
```

Figure \@ref(fig:occupation) shows that there are more women working in non-agricultural occupations. When we compare the mothers working in the agricultural sector, the number of women with a child under six is twice as many as the number of the ones without. On the contrary, the number of women with no child under six is slightly higher than that of the ones with a child in non-agricultural occupations.

```{r occupationchildcare, echo=FALSE, fig.cap="Distribution of child's caretaker by occupation", fig.width=6, fig.height=4}
# caretaker aspect
occup_caretaker <-
  occup_long %>%
  filter(str_detect(variable, "^caretaker_")) %>%
  mutate(variable = str_replace_all(variable, "caretaker_", "")) %>%
  rename(caretaker = variable,
         percentage = count) %>%
  mutate(
    caretaker = factor(caretaker, 
                       levels = c('respondent',
                                  'partner',
                                  'other_relative',
                                  'neighbor',
                                  'hired',
                                  'child_in_school',
                                  'other_female_child',
                                  'other_male_child',
                                  'not_worked_since_birth',
                                  'other',
                                  'missing'))
  )

# plot bar graph
occup_caretaker %>%
  ggplot(aes(x=caretaker, y=percentage, fill=background)) +
  geom_bar(stat = "identity", position="dodge") + 
  labs(x='Caretaker while at work', y="Percent", fill="Occupation") +
  theme_minimal() +
  theme(legend.position='bottom') +
  coord_flip() +
  scale_x_discrete(
    labels = 
      c("respondent" = "respondent", 
        "partner" = "partner",
        "other_relative" = "relative",
        "neighbor" = "neighbor",
        "hired" = "hired",
        "child_in_school" = "schooling",
        "other_female_child" = "other female child",
        "other_male_child" = "other male child",
        "not_worked_since_birth" = "not worked since birth",
        "other" = "other",
        "missing" = "missing")
  )
```

Figure \@ref(fig:occupationchildcare) illustrates that the percentage of the respondent herself being the child's caretaker is highest,followed by relative and schooling regardless of the type of jobs they have. Women who work in the agricultural sector are more likely to care for a child themselves than women in non-agricultural occupations. In particular, more than a half of the employed women working in the agricultural industry take care of their child themselves whereas around 45% of the women working in non-agricultural sector take care of their child themselves. Another observation is that female siblings are more likely to look after their younger siblings compared to male siblings.

## Employment status

```{r employmentstatus, echo=FALSE, message=FALSE, fig.cap="Employment status", fig.width=6, fig.height=4}
# select employment status
employment_status <-
  ghana %>%
  filter(character=="Employment_status") %>%
  mutate(
    no_child_under_six = round(no_child_under_six/100 * number_of_employed_women, 0),
    one_more_child_under_six = round(one_more_child_under_six/100 * number_of_employed_women, 0)
  )

# make dataset in long format
emply_long <-
  employment_status %>%
  select(-c(character, total)) %>%
  pivot_longer(!background, names_to = "variable", values_to = "count") %>%
  mutate(
    background = factor(background,
                        levels = c('All_year_full_week',
                                   'All_year_part_week',
                                   'Seasonal',
                                   'Occasional'))
  )

emply_long %>%
  filter(variable == 'no_child_under_six' | variable == 'one_more_child_under_six') %>%
  ggplot(aes(x = background, y = count, fill = variable)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Employment status", y = "Number of employed women", fill = "Child under six") +
  theme_minimal() +
  scale_fill_discrete(labels = c("No child under six", 
                                 "One or more child under six")) +
  scale_x_discrete(labels = c("All_year_full_week" = "All year, full week",
                              "All_year_part_week" = "All year, part week")) +
  theme(axis.text.x = element_text(angle=15), legend.position = 'bottom')
```

Figure \@ref(fig:employmentstatus) illustrates that majority of employed women work throughout the entire year. Specifically, majority of them work full week. There are few women who work occasionally. On top of that, regardless of their employment status, more working mothers tend to have one or more child under six years old.

```{r employmentstatuschildcare, echo=FALSE, fig.cap="Distribution of child's caretaker by work status", fig.width=6, fig.height=4}
# caretaker aspect
emply_caretaker <-
  emply_long %>%
  filter(str_detect(variable, "^caretaker_")) %>%
  mutate(variable = str_replace_all(variable, "caretaker_", "")) %>%
  rename(caretaker = variable,
         percentage = count) %>%
  mutate(
    caretaker = factor(caretaker, 
                       levels = c('respondent',
                                  'partner',
                                  'other_relative',
                                  'neighbor',
                                  'hired',
                                  'child_in_school',
                                  'other_female_child',
                                  'other_male_child',
                                  'not_worked_since_birth',
                                  'other',
                                  'missing')
    )
  )

# plot the bar graph
emply_caretaker %>%
  ggplot(mapping=aes(x=caretaker, y=percentage, fill=caretaker)) +
  geom_bar(stat = "identity", position="dodge") + 
  theme_minimal() +
  labs(x='Caretaker while at work', y="Percentage (in %)") +
  theme(legend.position='none') +
  coord_flip() + 
  scale_x_discrete(
    labels = 
      c("respondent" = "respondent",
        "partner" = "partner",
        "other_relative" = "relative",
        "neighbor" = "neighbor",
        "hired" = "hired",
        "child_in_school" = "schooling",
        "other_child" = "siblings",
        "not_worked_since_birth" = "not worked since birth",
        "other" = "other",
        "missing" = "missing")
  ) +
  facet_wrap(vars(background),
             dir = "v",
             scales = "free",
             labeller = labeller(background = c("All_year_full_week" = "All year, full week",
                                                "All_year_part_week" = "All year, part week",
                                                "Seasonal" = "Seasonal",
                                                "Occasional" = "Occasional")
                                 )
  )

```

We can see from Figure \@ref(fig:employmentstatuschildcare) that the proportion of respondents being the primary caretaker is remarkably higher than the other caretaker options in all four employment status. It exceeds 40% in all four cases, and in particular, mothers working full time are more likely to look after their own children than the others. Next, it is the relatives who mainly take care of the child under six. It should be noted that there are more women sending their kids under six to school when they work all year compare to the ones who do not work throughout the entire year.

# Discussion

## Employed Women

We can tell from Table \@ref(tab:data) that 3,564 out of 4,843 women of age 15-49 who completed the survey are employed and have children. In other words, around three in four women were employed during the 12 months before the survey. Figure \@ref(fig:residence) shows that the number of employed women is greater in rural area. It is worth mentioning that even though rural areas typically have low population density, rural women are somewhat more likely than urban women to be currently working. Next, Figure \@ref(fig:region) illustrates that women residing in the Upper West Region are least likely to be employed. Together with Figure \@ref(fig:map), we can tell that the women are more likely to be employed when they live in the southern part of Ghana. The northern part of Ghana including Upper West, Upper East, and Northern Region are the bottom 3 regions in Figure \@ref(fig:region) which demonstrates the distribution of the number of employed women whereas the southern part of Ghana including Greater Accra, Ashanti, and Eastern Region are the top 3.

Additionally, it is surprising to see from Figure \@ref(fig:motheredu) that the number of employed women whose education level is secondary school or higher is remarkably lower than that of the other education levels. Meanwhile, Figure \@ref(fig:occupation) shows that women are approximately twice as likely to be employed in the non-agricultural sector than in the agricultural sector. Lastly, when women are working, then Figure \@ref(fig:workstatus) tells that they are most likely to be self-employed and Figure \@ref(fig:employmentstatus) shows that they are most likely to work full-time.

## Presence of Child Under Six Years Old

Table \@ref(tab:data) shows that more than a half of the employed women has a child under six years. Figure \@ref(fig:childstatus) demonstrates that working women in rural area are more likely to have at least one child under six years whereas working women in urban area tend to have no child under six years. Figure \@ref(fig:motheredu) shows that as the level of working mother’s education gets higher, it is more likely to not have any child under six.

Figure \@ref(fig:workstatus) illustrates that when they work for themselves or for their family, then there are more women with a child under six whereas when they work for someone else, then there are more women without. Figure \@ref(occupation) shows that when women are working in the agricultural sector, they are more likely to have a child under six whereas when they work in the non-agricultural sector, it is equally likely to have a child under six and without. Lastly, Figure \@ref(fig:employmentstatus) tells that it is more likely to have a child under six years regardless of their employment status. In other words, no matter whether they work full-time or part-time, there are more women who have at least one child under six years.

## Child's Caretaker

Every figure that demonstrates the distribution of the child's caretaker shows that it is most likely to be the mother primarily taking care of their child under six while they are at work. In other words, no matter where they live, what their education level is, and what type of a job they have, majority of the employed women is the look after their child. Therefore, we can deduce that in Ghana it was common that the mother is primarily responsible for the care and raising of a child back in 1998.

Besides the mother being the primary caregiver of a child under six, it can be observed from Figure \@ref(fig:mothereduchildcare), Figure \@ref(fig:workstatuschildcare), Figure \@ref(fig:occupationchildcare), Figure \@ref(fig:employmentstatuschildcare) that the relatives other than husband/partner are likely to look after children under six. Also, there are some family where the older siblings take care of their siblings under six. Thus, we can tell that if there is a child under six in the family, then it was usual for the family member to take care of them in Ghana, 1998.

In addition, we can observe that it was uncommon to hire a person to take care of the child under six. Instead of hiring a person, there are more cases reported that they sent the kids to school or to institutional care. Figure \@ref(fig:areachildcare) demonstrates that the percentage of the employed women sending their kids to school is higher in urban areas than in rural areas. Figure \@ref(fig:mothereduchildcare) shows that the number of working mothers sending their kids to school increases as their mother's education level gets higher. Figure \@ref(fig:workstatuschildcare) illustrates that it is more likely to have their kids in school when they work for someone else than when they work for the family member or when they are self-employed. Figure \@ref(fig:occupationchildcare) shows that the percentage of the employed women sending their kids to school is higher in non-agricultural sectors than in agricultural sectors. Lastly, we can see from Figure \@ref(fig:employmentstatuschildcare) that there are more women sending their kids to school when they work all year compare to the ones who work seasonally and occasionally.

Lastly, we can tell that it is rarely the case that the mother stops working after having a birth. However, it is worth noting that the percentage of the employed women stop working after the birth is higher in urban areas than in rural areas from Figure \@ref(fig:areachildcare). In addition, Figure \@ref(fig:mothereduchildcare) shows that the women who completed high school have more tendency to stop working after having a birth than the ones with a lower education level. Furthermore, Figure \@ref(fig:employmentstatuschildcare) shows that the percentage of the employed women stop working after the birth is highest when they work all year, full week.

## Limitation and weaknesses

4,843 women of age 15-49 completed the survey, and 3,564 of them responded that they are employed. Then, they were further asked whether they have a child under six. However, it is not mentioned whether their marital status was asked. If the question was asked to every women who participated the survey, then the respondents who answered to have no child under six might be single. This dataset is created to identify woman's overall status, and the percentage of the employed mothers not having a child is one of the important indicators. However, as it is unclear which group that not having a child under six refers to, it is hard to tell what factor this survey focuses on.

Additionally, it would have been better if they were first asked for their marital status and the presence of dependent children. They could have then asked the age of youngest child in the household or whether they have a child under six in the household. As married life has been reported to have a gendered effect on women and men's labour market outcomes [@maritalstatus], the distribution of the marital status among the employed women would be useful to tell the woman's overall status. In this context, if the respondent answered that they stopped working since the last birth, they could have further asked the reason for stopping working and about the job that they used to have before having a birth.

Lastly, only the employed women who have a child under six were further asked to indicate who takes care of their children while they are at work. The dataset included the percentage of the missing values, and we purposely excluded in plotting the graphs as they do not tell anything about the data itself and there were only few missing values. However, when we add up the percentage of all possible items including the missing values, there are several background characteristics where the numbers did not add up to a 100. As only the percentage is provided in the report, there might be a round-off error. Particularly, since the numbers were rounded to one-decimal place, there is a high possibility that there was a rounding error. To make the dataset more precise, they could have rounded to two-decimal places.

On top of that, the survey asked for the employment situation during the 12-month reference period preceding the survey. However, the respondents were allowed to answer using their previous experiences. In particular, this applies to the respondents who had a job before having a birth but stopped working since the birth. Since we do not know the age of the youngest child in their household, they might have answered the questionnaire based on their employment situation prior to the 12-month reference period. Therefore, the time period does not match among the questions.

\newpage

# References